---
title: "STA 141A Final Project: Predictive Model of Mice Neural Activity"
author: "Nathaniel Faxon 917100572"
date: "2023-04-18"
output: html_document
abstract: |
  In this project, we were given 18 sessions of mice brain neural activity, each session recording brain activity related variables as well as the feedback produced from the mice after completing a visual discrimination task. With this information, data was combined into multiple dataframes and important variables were classified. Clustering and classification techniques were used in order to combine similar data across different trials (3 clusters used). Using Principal Component Analysis (PCA), the dataframe dimensions were reduced in order to capture more important features and get rid of noise. Logistic Regression was then applied to form our predictive model.
---

```{r}
# Load the relevant libraries

suppressWarnings(library(tidyverse))
suppressWarnings(library(ggplot2))
suppressWarnings(library(knitr))
suppressWarnings(library(dplyr))
suppressWarnings(library(gtools))
suppressWarnings(library(corrplot))
suppressWarnings(library(gridExtra))
suppressWarnings(library(grid))
suppressWarnings(library(PerformanceAnalytics))
suppressWarnings(library(grDevices))
suppressWarnings(library(viridis))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(reshape2))
suppressWarnings(library(ggcorrplot))
suppressWarnings(library(DT))
```

```{r}
# Load the session information

session=list()
for(i in 1:18){
  session[[i]]=readRDS(paste('./Data/session',i,'.rds',sep=''))
  #print(session[[i]]$mouse_name)
  #print(session[[i]]$date_exp)
  
}

```

## Introduction

```         
In the original study, the neural activity of 30,000 neurons in 42 brain regions of mice were analyzed. There were 10 mice analyzed over 39 different sessions. Within each session, visual stimuli were presented to the mice at random, then the reaction of the mice to this stimuli were recorded.

In this project, we will be analyzing the different trials and recorded neuron spikes in order to make a predictive model for feedback type. This research is important because it can help one understand the different brain regions that participate in action selection. Past studies about action selection tended to focus on individual brain region. This study could help explain how other regions like, sensory regions, could also participate in action selection. In the end, this study provides the organizing principles for the distribution of how neurons encode behavioral relevant variables in the brain of a mouse.

The data comes from 18 different session of 4 different mice brain neural activity. The four mice that we are studying the neural responses of are Cori, Forssmann, Hench, and Lederberg. More specifically, the first three sessions analyzed Cori, the next four Forssmann, the next four Hench, and the last seven Lederberg. These sessions were taken between 12/14/2016 through 12/11/2017.

Within each session, there were many trials where the activity of the neurons in the mice's visual cortex was recorded based on their decision making when faced with visual stimuli. Their decision making consists of turning the wheel either to the left or right or holding it still. Different variables were recorded for the 4 different mice. 8 variables are recorded each session:

-   brain_area: The area of the brain where each neuron lives.
-   contrast_left: Contrast on the left stimulus. Can take values 0, 0.25, 0.5, 1 (0 indicating the absence of a stimulus).
-   contrast_right: Contrast on the right stimulus. Can take values 0, 0.25, 0.5, 1 (0 indicating the absence of a stimulus).
-   date_exp: Date of the experiment on the mouse.
-   feedback_type: The type of feedback that was administered base on the outcome of the mouse's decision. There are four different cases that can happen:
    -   When the right contrast is greater than the left contrast: If the mouse turned the wheel to the left, feedback_type = 1. If turned to the right, feedback_type = -1.
    -   When the left contrast is greater than the right contrast: If the mouse turned the wheel to the right, feedback_type = 1. If turned to the left, feedback_type = -1.
    -   When the left contrast and right contrast equal 0: If mouse held the wheel still, feedback_type = 1. Otherwise, feedback_type = -1.
    -   When the left contrast and right contrast are equal and not both 0: There is a 50% chance that feedback_type = 1 or -1.
-   mouse_name: Name of the mouse being used in the experiment (trial). Specific to this project, the 4 mice are Cori, Forssmann, Hench, and Lederberg.
-   spks: Numbers of spikes of neurons in the visual cortex. Holds many different trials.
-   time: Center of the time bins for the spks.

From the existing research, neuron patterns within the mice have emerged when responding to their visual discrimination task. Specific areas within their brain are occupied when the neurons encode visual stimuli. But in completing the overall task, all brain areas had pre-stimulus activity with enhanced subcortical and suppressed neocortical activity during engagement.
```

## Exploratory Analysis

-   Basic descriptive statistics (missing, mean, quartiles, \# of observations)
-   Data visualization (features, patterns in data)

```{r}

# Get the variables recorded within this session
print(ls(session[[1]])) # or names(session[[1]]))

#Determine the different variables and details within each session

uniqueBrainareasVector <- c()

for (i in 1:18) {
  
  message("Session ", i)
  
  session_i <- session[[i]]
  
  print(session[[i]]$mouse_name)
  print(session[[i]]$date_exp)
  
  # Get the affected brain areas within this session
  print("Affected brain areas:")
  print(unique(session_i$brain_area))
  
  uniqueBrainareasVector <- append(uniqueBrainareasVector, unique(session_i$brain_area))
  
  # Get the number of trials within the session
  print("Number of trials:")
  print(length(session_i$feedback_type))
  
}

# Get the unique brainareas through all sessions
uniqueBrainareasVector <- unique(uniqueBrainareasVector)
uniqueBrainareasVectorLength = length(uniqueBrainareasVector)
message("There are ", uniqueBrainareasVectorLength, " unique brain areas throughout all sessions:")
print(uniqueBrainareasVector)



```

```{r}
# Dataframe for the average spks per trial for each session

# Function to calculate the average spks per brain area for the specific trial of a session
firing_rate_brainarea <- function(trialNum, currSession) {
  
  spkTrial = currSession$spks[[trialNum]]
  currBrainArea = currSession$brain_area
  spkCount = apply(spkTrial, 1, sum)
  spkAvg = tapply(spkCount, currBrainArea, mean)
  return(spkAvg)
  
}


# Function to find average spks per brain area for all trials in a session
trial_firing_rate_brainarea <- function(sessionNum) {
  
  trialLen = length(session[[sessionNum]]$feedback_type)
  brainareaLen = length(unique(session[[sessionNum]]$brain_area))
  
  trialBrainareaSummary = matrix(nrow = trialLen, ncol = brainareaLen + 2)
  
  for(q in 1:trialLen) {
    
    trialBrainareaSummary[q,] = c(firing_rate_brainarea(q, currSession = session[[sessionNum]]), sessionNum, q)
    
  }
  
  colnames(trialBrainareaSummary) = c(names(firing_rate_brainarea(q,currSession = session[[sessionNum]])), 'session_number', 'trial_number')
  
  trialBrainareaSummary <- as_tibble(trialBrainareaSummary)
  
  
  # Plot the spikes per area in each session against trials
  colorPalette = brewer.pal(n=brainareaLen, name = 'Paired')

  plot(0~1, col='white',xlim=c(0,trialLen),ylim=c(0.5,2.2), xlab="Trials",ylab="Average Spike Counts", main=paste("Spikes per Area in Session", sessionNum))
  
  for(i in 1:brainareaLen){
    lines(y=trialBrainareaSummary[[i]],x=trialBrainareaSummary$trial_number,col=colorPalette[i],lty=2,lwd=1)
    lines(smooth.spline(trialBrainareaSummary$trial_number, trialBrainareaSummary[[i]]),col=colorPalette[i],lwd=3)
  }
  
  legend("topright", 
         legend = colnames(trialBrainareaSummary)[1:brainareaLen], 
         col = colorPalette, 
         lty = 1, 
         cex = 0.8
  )
    
    
    # Plot the correlation matrix for the brain areas, see if there are relations between brain areas
    correlationMatrixPlot <- suppressWarnings(chart.Correlation(trialBrainareaSummary[, 3:brainareaLen], histogram = TRUE, pch = 7, alpha = 0.3))
    correlationMatrixPlot
    
  
  # Return the dataframe of brain areas for one session
  return(trialBrainareaSummary)
  
}

# Use above function to apply to all sessions, bind them together to make one dataframe of all session brainarea spk averages. 

brainarea_firing_rateDF <- data.frame()

for (p in 1:18) {
  
  brainarea_firing_rateDF = bind_rows(brainarea_firing_rateDF, trial_firing_rate_brainarea(p))

}

brainarea_firing_rateDF <- relocate(brainarea_firing_rateDF, trial_number)
brainarea_firing_rateDF <- relocate(brainarea_firing_rateDF, session_number)

#head(brainarea_firing_rateDF)
#datatable(brainarea_firing_rateDF)

# In the correlation matrices:
#The distribution of each variable is shown on the diagonal.
#On the bottom of the diagonal : the bivariate scatter plots with a fitted line are displayed
#On the top of the diagonal : the value of the correlation plus the significance level as stars
#Each significance level is associated to a symbol : p-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(“***”, “**”, “*”, “.”, " “)

```

```{r}
# Dataframe with general statistics across sessions

sessionDF <- data.frame()

for (i in 1:18) {

  #Create a custom dataframe here with 4 columns for 4 variables
  trialDF <- data.frame(matrix(ncol = 5, nrow = 0))

  for (j in 1:length(session[[i]]$spks)) { 
    
    # Calculate the total number of neurons active
    neuronCountTable = table(unlist(session[[i]]$spks[[j]]))
    neuronCountValues = as.numeric(neuronCountTable)
    numActiveNeurons = length(session[[i]]$spks[[j]]) - neuronCountValues[1] #neuronCountValues[1] is the # of 0's
    
    # Calculate the total number of spks
    numSpks = 0
    for(k in 1:length(neuronCountTable)) {
    
      numSpks = numSpks + ((k-1) * neuronCountValues[k])
      
    }
    
    # Calculate the proportion of active neurons
    activeNeuronProp = numActiveNeurons / length(session[[i]]$spks[[i]])
    
    # Calculate the firing rate (total number of spks != 0 divided by total)
    firing_rate = numSpks / (numSpks + neuronCountValues[1])
    
    # Calculate the absolute difference between left and right contrast
    contrastDifference = abs(session[[i]]$contrast_left[j] - session[[i]]$contrast_right[j])
    
    # Append to the variables to the trial data frame
    trialDF[nrow(trialDF) + 1,] = c(contrastDifference, numActiveNeurons, numSpks, activeNeuronProp, firing_rate)
    
  }
  

  # Append data together using cbind
  tempSession = cbind(rep(i,length(session[[i]]$contrast_left)),
                       seq(1, length(session[[i]]$spks)),
                       session[[i]]$mouse_name, 
                       session[[i]]$contrast_left,
                       session[[i]]$contrast_right,
                       length(session[[i]]$brain_area),
                       length(unique(session[[i]]$brain_area)),
                       length(session[[i]]$spks),
                       trialDF,
                       session[[i]]$feedback_type
                       )
  
  # Append each session using rbind
  if (i == 1){
    sessionDF <- tempSession
  }
  else {
    sessionDF = rbind(sessionDF, tempSession)
  }
  
  
}


colnames(sessionDF) = c("session_number", "trial_number", "mouse_name", "contrast_left", "contrast_right", "total_number_of_neurons", "number_of_unique_brain_area", "number_of_trials", "contrast_difference", "number_of_active_neurons", "total_number_of_spks", "proportion_of_active_neurons", "firing_rate", "feedback_type")
sessionDF = as.data.frame(sessionDF)
sessionDF$session_number = as.factor(sessionDF$session_number)

#head(sessionDF)
#datatable(sessionDF)

```

```{r}
# Box plots of firing rate (or numberofactiveneurons, total num of spikes, or proportion) on feedbacktype
# Mention how spks and active neurons are different, since there can be multiple spks per neuron


# Box Plots


# Firing Rate on Feedback Type
firingRate_by_feedback_boxPlot <- ggplot(sessionDF, aes(x = as.factor(feedback_type), y = firing_rate)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Feedback Type") + ylab("Firing Rate") + ggtitle("Box Plot of Firing Rate on Feedback Type") + theme(legend.position="none")

# Number of Active Neurons on Feedback Type
numActiveNeurons_by_feedback_boxPlot <- ggplot(sessionDF, aes(x = as.factor(feedback_type), y = number_of_active_neurons)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Feedback Type") + ylab("Number of Active Neurons") + ggtitle("Box Plot of Number of Active Neurons on Feedback Type") + theme(legend.position="none")

# Total Number of Spikes on Feedback Type
totalNumSpikes_by_feedback_boxPlot <- ggplot(sessionDF, aes(x = as.factor(feedback_type), y = total_number_of_spks)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Feedback Type") + ylab("Total Number of Spikes") + ggtitle("Box Plot of Total Number of Spikes on Feedback Type") + theme(legend.position="none")

# Proportion of Active Neurons on Feedback Type
proportionActiveNeurons_by_feedback_boxPlot <- ggplot(sessionDF, aes(x = as.factor(feedback_type), y = proportion_of_active_neurons)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Feedback Type") + ylab("Proportion of Active Neurons") + ggtitle("Box Plot of Proportion of Active Neurons on Feedback Type") + theme(legend.position="none")


grid.arrange(firingRate_by_feedback_boxPlot, proportionActiveNeurons_by_feedback_boxPlot, numActiveNeurons_by_feedback_boxPlot, totalNumSpikes_by_feedback_boxPlot, nrow = 2, ncol = 2)


```

```{r}

# Correlation Matrices for each Session
for (i in 1:18) {
  
  corrplot.mixed(round(cor(sessionDF[sessionDF$session_number == i, c('contrast_left', 'contrast_right', 'contrast_difference', 'number_of_active_neurons', 'total_number_of_spks', 'proportion_of_active_neurons', 'firing_rate', 'feedback_type')]), 2), title = paste("Correlation Matrix of Session", i), mar=c(1,1,1,1))
  
}
```

```{r}

# Correlation Matrices for Each Mouse
mouseNameVector <- c('Cori', 'Forssmann', 'Hench', 'Lederberg')
for (i in 1:length(mouseNameVector)) {
  
  corrplot.mixed(round(cor(sessionDF[sessionDF$mouse_name == mouseNameVector[i], c('contrast_left', 'contrast_right', 'number_of_active_neurons', 'total_number_of_spks', 'proportion_of_active_neurons', 'firing_rate', 'feedback_type')]), 2), title = paste("Correlation Matrix of Mouse", mouseNameVector[i]), mar=c(1,1,1,1))
  
}

```

```{r}

# Plots for variables against trials (for individual session analysis)

for (i in 1:18) {
 
  plot1 <- ggplot(sessionDF[sessionDF$session_number == i,], aes(x = trial_number, y = number_of_active_neurons)) + geom_point() + geom_smooth(formula = y ~ x, method="lm") + ggtitle("Number of Active Neurons Against Trials") + labs(y = "Number of Active Neurons", x = "Trials")
  
  plot2 <- ggplot(sessionDF[sessionDF$session_number == i,], aes(x = trial_number, y = total_number_of_spks)) + geom_point() + geom_smooth(formula = y ~ x, method="lm") + ggtitle("Total Number of Spikes Against Trials") + labs(y = "Total Number of Spikes", x = "Trials")
  
  plot3 <- ggplot(sessionDF[sessionDF$session_number == i,], aes(x = trial_number, y = proportion_of_active_neurons)) + geom_point() + geom_smooth(formula = y ~ x, method="lm") + ggtitle("Proportion of Active Neurons Against Trials") + labs(y = "Proportion of Active Neurons", x = "Trials")
  
  plot4 <- ggplot(sessionDF[sessionDF$session_number == i,], aes(x = trial_number, y = firing_rate)) + geom_point() + geom_smooth(formula = y ~ x, method="lm") + ggtitle("Firing Rate Against Trials") + labs(y = "Firing Rate", x = "Trials")
  
  grid.arrange(plot1, plot2, plot3, plot4, nrow = 2, ncol = 2, top = textGrob(paste("Session", i, "Scatterplots")))
   
}

```

```{r}
# Plots for variables against trials (for session analysis)

ggplot(sessionDF, aes(x = trial_number, y = number_of_active_neurons, colour = session_number)) + geom_point() + geom_smooth(method = "loess", formula = y ~ x) + ggtitle("Scatterplot of Number of Active Neurons Against Trials") + labs(y = "Number of Active Neurons", x = "Trials")

ggplot(sessionDF, aes(x = trial_number, y = total_number_of_spks, colour = session_number)) + geom_point() + geom_smooth(method = "loess", formula = y ~ x) + ggtitle("Scatterplot of Total Number of Spikes Against Trials") + labs(y = "Total Number of Spikes", x = "Trials")

ggplot(sessionDF, aes(x = trial_number, y = proportion_of_active_neurons, colour = session_number)) + geom_point() + geom_smooth(method = "loess", formula = y ~ x)  + ggtitle("Scatterplot of Proportion of Active Neurons Against Trials") + labs(y = "Proportion of Active Neurons", x = "Trials")

ggplot(sessionDF, aes(x = trial_number, y = firing_rate, colour = session_number)) + geom_point() + geom_smooth(method = "loess", formula = y ~ x)  + ggtitle("Scatterplot of Firing Rate Against Trials") + labs(y = "Firing Rate", x = "Trials")


```

```{r}
# Combine both dataframes to one large one

completeDF <- cbind(sessionDF, subset(brainarea_firing_rateDF, select = -c(session_number, trial_number) ))
#head(completeDF)
datatable(completeDF)

```

```{r}
# Scatterplot for Mice Brain areas on feedback type per mouse

for (i in 1:18) {
  
  tempBrainAreaDF <- completeDF[completeDF$session_number == i, colSums(is.na(completeDF[completeDF$session_number == i,])) == 0]
  tempBrainAreaDF <- melt(tempBrainAreaDF, id.vars = "feedback_type", measure.vars = colnames(tempBrainAreaDF)[14:length(colnames(tempBrainAreaDF))])
  
  
  tempPlot <- ggplot(tempBrainAreaDF) + geom_boxplot(aes(x = as.factor(feedback_type), y = value, color = variable), outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Feedback Type") + ylab("Firing Rate") + ggtitle(paste("Session", i, "Box Plot of Brain Area Firing Rate on Feedback Type")) + guides(color = guide_legend(title = "Brain Areas"))
  
  print(tempPlot)
  
}

```

```{r}
# Scatterplot of mice brainarea firing rate by trial per mouse

tempBrainAreaPerMouse <- completeDF[completeDF$mouse_name == mouseNameVector[4], ]
tempBrainAreaPerMouse <- tempBrainAreaPerMouse[,colSums(is.na(tempBrainAreaPerMouse))<nrow(tempBrainAreaPerMouse)]
tempBrainAreaPerMouse <- tempBrainAreaPerMouse[, 14:length(colnames(tempBrainAreaPerMouse))]
tempBrainAreaPerMouse[is.na(tempBrainAreaPerMouse)] <- 0

ggplot(melt(tempBrainAreaPerMouse), aes(x = variable, y = value)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2, alpha = 0.2) + xlab("Brain Areas") + ylab("Firing Rate") + ggtitle(paste("Mouse", mouseNameVector[4], "Box Plots of Brain Area Firing Rates"))


```

## Methodology

-   Data Integration: PCA, clustering (improving prediction performance)
-   Predictive Model (GLM, linear regression, time series)

```{r}
# Perform KMeans clustering

completeDF[is.na(completeDF)] <- 0

# Clustering data for 3 clusters
completeDF.k3 <- subset(completeDF, select = -c(session_number, trial_number, mouse_name, feedback_type)) %>% kmeans(3)
clusterIndicatorVector <- completeDF.k3$cluster

# Append the cluster indicator to the dataframes for further analysis
completeDF <- cbind(clusterIndicatorVector, completeDF)
colnames(completeDF)[1] <- "cluster_id"

```




```{r}
#TODO: Ask if I should include the analysis on brain area (brain_areaDF, etc.) since I do not use the variables 
# Use discussion 10 format for the final project format, also shows how to show figures without code
# Explain variables used in dataframes that are printed



completeDF.pca <- subset(completeDF, select = -c(session_number, trial_number, mouse_name, feedback_type)) %>% prcomp(center = TRUE, scale = TRUE)

completeDF.pca$rotation
summary(completeDF.pca)

plot(completeDF.pca, type = 'l', main = "CompleteDF Scree Plot")


# TODO STEPS
# scale the entire completeDF
# Use logisitic regression on scaled completeDF

```


```{r}

# USe logisitic regression with the variables chosen above to form a prediction model
# Ask how to use the weighted variables from PCA

# First split the data into training and testing groups
completeDF_test <- slice(completeDF, 1:10)


completeDF.logReg <- glm(as.factor(feedback_type) ~ cluster_id + contrast_left + contrast_right + total_number_of_neurons + number_of_unique_brain_area + number_of_trials + contrast_difference + number_of_active_neurons, data = completeDF, family = "binomial")

summary(completeDF.logReg)

logRegpredictedModel <- predict(completeDF.logReg, completeDF_test, type = "response")

logRegpredictedModel.cm <- confusionMatrix(logRegpredictedModel$class, completeDF_test$feedback_type)
logRegpredictedModel.cm

paste("The misclassification error rate is 1 - accuracy =", 1 - logRegpredictedModel.cm$overall['Accuracy'])

```

5)  Result (prediction performance)

6)  Conclusion, Discussion

-   Findings, variable information, drawbacks/disadvantages, next steps

7)  References

# Extra

```{r}

dat <- session[[1]]

#session[[1]]$brain_area

unique(dat$brain_area)

i_trial = 12;
i_session = 1;

dim(session[[i_session]]$spks[[i_trial]])

length(session[[i_session]]$feedback_type)

apply(session[[i_session]]$spks[[i_trial]], 2, sum)

names(session[[2]])

#Use different spks trials to make predictive models (logistic regression will work, show some linear or svm)
#session[[i_session]]$spks[[i_trial]]
#to predict:
#session[[i_session]]$feedback_type[i_trial]

# LOOK AT COURSE PROJECT DESCRIPTION .RMD FOR VARIABLE EXPLORATION

#Ask if we are using the trials of spks in combination with other recorded variables (eg. brain_area, contrast_left, contrast_right) to predict the feedback_type. Yes

#Predict the feedback_type per trial, not session

# Create df including all sessions, each session df will have rows as trials, and the columns will be variables of each trial (contrast_left, contrast_right, firing_rate, feedback_type). Use bind to stack each session df.

```
